---
title: "Using Postgre with R"
output: html_notebook
---

First, we need to establish a connection with our database. For this, we will use the RPostgres and DBI libraries.

If you are working with a large database, you might consider using the dbplyr library (https://dbplyr.tidyverse.org/articles/dbplyr.html), which allows you to use dplyr verbs to interact with tables in a database, but actually execute them in the database.

```{r}
library(DBI)
library(RPostgres)
library(tidyverse)
```

```{r}
complete_data <- read_csv('../data/completed_table.csv')
```











Now, we'll create our connection.
```{r}
con <- dbConnect(Postgres(),                 # Use the postgres driver
                 dbname = 'metalworkingsolutions',     # Name of the database we want to connect to
                 host = 'localhost',         # 'localhost' or eg. 'ec2-54-83-201-96.compute-1.amazonaws.com'
                 port = 5432, 
                 user = 'postgres',
                 password = rstudioapi::askForPassword("Database password"))
```

We can see the tables in the database.
```{r}
dbListTables(con)
```

We can even see the columns of a particular table.
```{r}
dbListFields(con, 'parts')
```

To execute a query, we can use the dbSendQuery function
```{r}
query = "WITH parts_jobs AS(
SELECT
	imp_short_description
	,jmp_part_id
	,COUNT(jmp_job_id) AS n_of_jobs
FROM jobs a
FULL JOIN parts b
	ON a.jmp_part_id = b.imp_part_id
	GROUP BY 1,2),
job_ops_count AS(
SELECT 
	jmo_job_id
	,SUM(a.jmo_actual_production_hours) AS twenty_four
	,SUM(b.jmo_actual_production_hours) AS twenty_three
	,(SUM(a.jmo_actual_production_hours) + SUM(b.jmo_actual_production_hours)) AS total_hours
FROM job_operations_2024 a
FULL JOIN job_operations_2023 b
	USING(jmo_job_id)
GROUP BY jmo_job_id
ORDER BY twenty_four DESC
),
full_hours AS(
SELECT
	jmo_job_id
	,COALESCE(total_hours, twenty_three, twenty_four) AS hours
FROM job_ops_count
ORDER BY 2 DESC
),
join_cte AS(
SELECT
	jmo_job_id
	,hours
	,b.jmp_part_id
	,c.imp_short_description
	,n_of_jobs
FROM full_hours a
JOIN jobs b
	ON a.jmo_job_id = b.jmp_job_id
JOIN parts c
	ON b.jmp_part_id = c.imp_part_id
JOIN parts_jobs d
	ON c.imp_part_id = d.jmp_part_id
ORDER BY 2 DESC)
SELECT 
	imp_short_description
	,jmp_part_id
	,SUM(hours) AS production_hours
	,n_of_jobs
FROM join_cte
GROUP BY imp_short_description, jmp_part_id, n_of_jobs
ORDER BY 3 DESC;"

res <- dbSendQuery(con, query)
```

Then to retrieve the results, we can use dbFetch. Note that you can fetch only a portion of the results if needed, but we'll fetch everything.
```{r}
parts_production <- dbFetch(res)

# It is good practice to clear the result after fetching what you need in order to free all resources associated with the result set. 
dbClearResult(res)
```

The results that are returned come through as a dataframe, so we can manipulate it using tidyverse.
```{r}
parts_production %>% 
  filter(n_of_jobs > 50)
```